---
title: "Week 1, Exercise 2: IPUMS EDA"
output:
  word_document: default
  pdf_document:
    includes:
      in_header: header_noR.txt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set Up

Download the `.dta` files for data extracts fro Armenia 2011, Costa Rica 2011 and Uruguay 2011

Read them into R using the haven package

```{r, message=FALSE,echo=TRUE, eval=FALSE}
library(tidyverse)
library(haven)

d1 <- read_dta("#####")
d2 <- #####
d3 <- #####
```

For each sample select only the sample, perwt, age, sex and chborn columns

```{r, echo=TRUE, eval=FALSE}
d1 <- select(d1, #####, #####, #####, #####, #####)
##### <- #####(d2, #####, #####, #####, #####, #####)
##### <- #####(d3, #####, #####, #####, #####, #####)
```

Use the `as_factor()` function to convert the sample column from labels to factors

```{r, echo=TRUE, eval=FALSE}
d1$##### <- #####(d1$sample)
#####$sample <- #####(#####$#####)
#####$##### <- #####(#####$#####)
```

Combine each data set into on single data frame

```{r, warning=FALSE, echo=TRUE, eval=FALSE}
d0 <- bind_rows(d1, #####, #####)
```

Use the `droplevels()` function to drop unused labels

```{r, echo=TRUE, eval=FALSE}
d0 <- #####(d0)
```


## Dimensions

How many observations are there in each sample?
What percent of indviduals are from each sample?

```{r, echo=TRUE, eval=FALSE}
d0 %>% 
  group_by(#####) %>%
  #####(n = #####(), 
            percent = #####()/nrow(d0) * 100)
```


## Summary Statistics 

For each of the countries in your dataset, we want to find 
    1) Mean 
    2) Median 
of the total number of children ever born [`CHBORN`]

Before making the calulations consider:

Q: What are the values of the `CHBORN` variable? What are the maximum number of children born to a woman
A: , 
```{r, echo=TRUE, eval=FALSE}
table(d0$#####)
```


Q: There are missing values in `CHBORN`. Before calculating the mean and median, you need to address the missing values. What is the difference between the two missing-value categories 98: Unknown and 99: NIU (not in universe)?
A: 



Q: Should you include or exclude cases with the missing values from your calculations?
A: 




Q: What is the population universe for this variable in each country? That is, in each census, who was asked this question?
A: 


Q: To make an accurate comparison of children ever born across these three countries, which cases should you drop from your dataset?
A: 

Q: Given the above, use `dplyr` functions to calculate the mean, median and sample size for each country:
A: 
```{r, echo=TRUE, eval=FALSE}
d0 %>% 
  #####(age >= 15, 
         sex == #####,
         chborn != #####,
         ##### != 99) %>%
  #####(sample) %>%
  summarise(mean = #####(chborn), 
            median = median(#####), 
            n = n())

```

Q: Calculate the lower quartile, upper quartile, interquartile range [IQR] and standard deviation of the total number of children ever born.

A: 
```{r, echo=TRUE, eval=FALSE}
d0 %>% 
  filter(##### >= 15, 
         ##### == 2,
         ##### != 98,
         chborn != 99) %>%
  group_by(#####) %>%
  summarise(q1 = quantile(chborn, probs = 0.25), 
            q3 = #####(chborn, probs = 0.75),
            iqr = #####(chborn),
            sd = #####(#####))
```

Q: Create a bar chart that displays the mean number of children born to women age 15 and older by country
A: 
```{r, echo=TRUE, eval=FALSE}
d4 <- ##### %>% 
  filter(##### >= 15, 
         sex == #####,
         chborn != #####,
         chborn != 99) %>%
  #####(sample) %>%
  #####(mean = mean(chborn))

ggplot(data = d4, 
       mapping = aes(x = #####, ##### = mean)) +
  geom_bar(stat = "identity")
```

The mean for `CHBORN` is calculated for women 15 & above. Recalculate the mean limiting your data to: (1) Women age 21 and older (2) Women age 30 and older and (3) Women above 45 and older:

```{r, echo=TRUE, eval=FALSE}
get_chborn <- function(d, x = 15){
  d %>% 
    filter(age >= x, 
           sex == 2,
           chborn != #####,
           chborn != 99) %>%
    group_by(sample) %>%
    summarise(mean = mean(chborn)) %>%
    mutate(age = x)
}

get_chborn(d0, x = 15)
get_chborn(#####, x = 21)
get_chborn(d0, x = 30)
get_chborn(d0, x = #####)
```

Create a boxplot (A) representing the number of children born (CHBORN) to women age 15 and older by country

```{r, echo=TRUE, eval=FALSE}
ggplot(data = d0 %>% 
  #####(age >= 15, 
         sex == #####,
         chborn != #####,
         chborn != 99),
       mapping = aes(x = #####, y = chborn)) +
  #####_boxplot()
```

Create a boxplot (B) representing the number of children born (CHBORN) to women age 15 and older by country

```{r, echo=TRUE, eval=FALSE}
ggplot(data = d0 %>% 
  filter(age >= #####, 
         ##### == 2,
         chborn != 98,
         chborn != 99),
       mapping = aes(x = sample, y = chborn)) +
  geom_#####()
```
Based on both box plots above, which of the following is true?

A) In box plot B, there is no median for Armenia because there not enough outliers.
B) In box plot B, the median in Armenia is the smallest since it has the least dots on top of its upper whisker.
C) In box plot A, the range is equal in all three countries
D) In box plot A, the IQR in Uruguay is the smallest, since its median is the smallest
