---
title: 'Week 2 Exercise 1: IPUMS Distributions'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### A. Set Up

1. Load the `haven`, `tidyverse` and `ipumsr` packages

```{r, message=FALSE}
library(haven)
library(tidyverse)
library(ipumsr)

```

2. Read in the `ipumsi_pan1960.dta` file, save as object `d0`

```{r}
d0 <- read_dta("C:/Users/18800/Desktop/stat model/ipumsi_pan1960.dta")
d0
```

#### Finding Married Couples in Household

3. Print household 1000 (`serial == 1000`)

```{r}
filter(d0, serial == 1000)
```

Q: How many people live in this house?

A: 10

Q: How many married couples are there? Who is married to who? Hint: inspect the pernum (person number) and sploc (spouse location) columns
    
A: 2 couples 
  1 married to 2
  3 married to 4

4. Create object `d1` that 
d1<- d1%>%
  filter(sploc!=0)%>%
  select(serial, age, `sex`, `pernum`, `sploc`)
    a. contains only those with a spouse in the household
    b. contains only the `serial`, `age`, `sex`, `pernum`, `sploc` columns
    c. `age` is numeric
    d. contains a new variable that uniquely identifies a married couple in a household
    e. arranged by household number, couple id and gender
   
```{r}
d1 <- d0 %>%
  filter(sploc != 0) %>%
  select(serial,age,sex, pernum, sploc) %>%
  mutate(age = as.numeric(age), 
         couple_id = paste(pmin(pernum, sploc), pmax(pernum, sploc), sep = "-")) %>%
  arrange(serial, couple_id, sex)
d1
```

#### Martial Age Differences

5. Modify `d1` so that it 

    a. contains the age difference of every couple in every household
    b. the husband person number
    c. the wife person number

```{r}
d1 <- d1 %>%
  group_by(serial, couple_id) %>%
  summarise(age_diff = age[1] - age[2],
            husband = pernum[1], 
            wife = pernum[2])
d1

```

6. Plot a density plot of the age differences, include a vertical line for the mean age difference in `d1`

```{r}
ggplot(data = d1, 
       mapping = aes(x = age_diff)) +
  geom_density() +
  geom_vline(xintercept = mean(d1$age_diff))
```

#### Z-Scores

7. Using the age differences in the Panama sample, what is the Z-score of 

    a. a marrital age difference of +10 years (male is 10 years older than female)
    b. a marrital age difference of -10 years (female is 10 years older than male)

```{r}
z_score <- function(x, mu, sigma){
  (x - mu)/sigma
}
m_age_diff <- mean(d1$age_diff)
s_age_diff <- sd(d1$age_diff)
z_score(x = 10, mu = m_age_diff, sigma = s_age_diff)
z_score(x = -10, mu = m_age_diff, sigma = s_age_diff)
```

8. For each of these cases, explain in your own words what the Z-scores mean.

    a. a case of marrital age difference of +10 years is 0.5417958 unit away from zero, we can see from the density plot that it's very rare

    b. a case of marrital age difference of -10 years is -1.961454 unit away from zero, we can see from the density plot that it's very rare

9. Calculate the Z-score for all couples in d1

```{r}
d1 <- d1 %>%
  mutate(age_diff_z = z_score(x = age_diff , mu = m_age_diff, sigma = s_age_diff))
d1
```

10. Plot a density plot of the Z-score of the age differences. 

```{r}
ggplot(data = d1, 
       mapping = aes(x = age_diff_z)) +
  geom_density()
```

11. Q: How does the plot differ from the previous density plot

    A: looks more normally distributed

#### Normality

12. Asses the closeness of the sample of age differences in d1 to Normal distribution using the `geom_qq()` and `geom_qq_line()` functions

```{r}
ggplot(data = d1, 
       mapping = aes(sample = age_diff_z)) +
  geom_qq() +
  geom_qq_line()
```

13. Add information on the wife's education by 

    a. filtering from `d0` the females with a spouse 
    b. selecting from `d0` the serial, pernum and edattain columns
    c. converting the `NIU` and `unknown` values of edattian to `NA`, 
    d. converting `edattain` to a factor
    e. renaming the `pernum` column to `wife`, ready for the join...
    f. using `right_join()` function to attach the wife education data to `d1`
    g. drop rows with `NA` for the wife `edattain`

```{r}
head(d0)
d2 <- d0 %>%
  filter(sex == 2, sploc != 0) %>%
  select(serial, pernum, edattain) %>%
  mutate(edattain = lbl_na_if(edattain, ~.val %in% c(0, 9)),
         edattain = as_factor(edattain)) %>%
  rename(wife = pernum) %>%
  right_join(d1) %>%
  drop_na()
```

14. Create a facet of qq plots by the wife's `edattain` status. 

```{r}
ggplot(data = d2, 
       mapping = aes(sample = age_diff_z)) +
  facet_wrap("edattain") +
  geom_qq() +
  geom_qq_line()
```

15. Q: What can we interpret from the plots?
    
    A: #####
     
       #####
       
       #####
       
       
#### Shapiro Test

16. Use the `shapiro.test()` function to test for normality from a sample of the age differnces. 

```{r}
s0 <- d1 %>%
  pull(age_diff) %>%
  sample(size = 5000)
shapiro.test(x = s0)
```

17. Interpet the output?

A: #####

18. Use the `shapiro.test()` function to test for normality from a sample of the age differnces from those where the wives completed university education. 

```{r}
s1 <- d2 %>%
  filter(edattain == "university completed") %>%
  pull(age_diff)
shapiro.test(x = s1)
```

19. Interpet the output?

A: #####