---
title: "Week 2 Exercise 1: IPUMS Distributions"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### A. Set Up

1. Load the `haven`, `tidyverse` and `ipumsr` packages

```{r, message=FALSE}
library(haven)
library(tidyverse)
library(ipumsr)
```

2. Read in the `ipumsi_pan1960.dta` file, save as object `d0`

```{r}
d0 <- read_dta("C:/Users/18800/Desktop/stat model/data/ipumsi_pan1960.dta")
d0
```

#### Confidence Intervals

3. Examine labels for the `chborn` variable

```{r}
table(d0$chborn)
```

4. Create an object `d1` that exlcudes all males and females with unknown (98) or not in universe (99) for `chborn`
 
```{r}
d1 <- d0 %>%
  filter(chborn != 98, 
         chborn != 99)  

```

5. Find the 95% confidence interval for the mean of `chborn` in Panama using the `t.test()` function

```{r}
t95 <- t.test(x = d1$chborn)
t95
t95$conf.int
```

6. Find the 80% confidence interval for the mean of `chborn` in Panama using the `t.test()` function

```{r}
t80 <- t.test(x = d1$chborn, conf.level = 0.8)
t80$conf.int
```


#### Proportions

7. Inspect the `edattain` codes

```{r}
str(d0$edattain)
table(d0$edattain)
unique(d0$edattain)
```

8. Create and object `tab0` that 

a. converts those with unknown (9) sex to `NA`
b. converts those not in universe (code 0) or unknown (9) to `NA`
c. converts sex and edattain to factors
d. creates a cross tabulation by sex and edattain

```{r}
tab0 <- d0 %>%
  mutate(sex = lbl_na_if(sex, ~.val == 9),
         edattain = lbl_na_if(edattain, ~.val %in% c(0,9)),
         sex = as_factor(sex),
         edattain = as_factor(edattain)) %>%
  select(sex, edattain) %>%
  table()
tab0
```

9. Display `tab0` with the margin totals

```{r}
addmargins(tab0)
```

10. Display the cross tabulation as a proportion table rounded to 3 decimal places
```{r}
tab0 %>% 
  prop.table() %>%
  round(digits = 3)
```

11. Use the `margin` argument in `prop.table()` to give the proportion by gender in each eduaction category rounded to 3 decimal places
```{r}
tab0 %>% 
  prop.table(margin = 2) %>%
  round(digits = 3)
```

12. Find a confidence interval for proportion of females with completed university education
```{r}
p0 <- prop.test(x = 100, n = 319)
p0
```

### Plotting Confidence Intervals

13. Create an object `d2` that exlcudes 
a. includes only the first person in the household (`pernum == 1`), 
b. excludes those with the number of rooms unknown (98) or not in universe (99)
c. excludes those with the martial status unknown (0) or not in universe (9)
d. converts `marst` to a factor
```{r}
d2 <- d0 %>% 
  filter(pernum == 1,
         marst != 0, 
         marst != 9, 
         rooms != 98,
         rooms != 99) %>%
  mutate(marst = as.factor(marst))
```

14. Modify `d2` so that is summarises for each martial status
a. the number of households,
b. the mean number of rooms
c. the standard error of the number of rooms
d. the upper and lower 95% bounds for a t-test for the number of rooms - calculate directly using `qt` (see page 18 in Gelman and Hill)

```{r}
d2 <- d2 %>%
  group_by(marst) %>%
  summarise(n = n(),
            mu = mean(rooms),
            se = sd(rooms)/sqrt(n),
            lwr = mu + qt(p = 0.025, df = n-1) * se,
            upp = mu + qt(p = 0.975, df = n-1) * se)
d2
```

15. Plot the confidence intervals for each martial status
```{r}
ggplot(data = d2, 
       mapping = aes(x = marst, y =mu)) +
  geom_point() +
  geom_pointrange(mapping = aes(ymin = lwr, ymax = upp))
```

16. Q: What can we say about the confidence intervals?

A: #####

17. Create an object `d3` based on `d0`
a. that excludes those with the eduacation unknown (0) or not in universe (9)
b. converts `edattain` to a factor
c. summarises by education category the
    i. number of males
    ii. number of females
    iii. number of people
    iv. the proportion of females in each group
    v. the standard error for the proportion
    vi. the upper and lower 95% confidence intervals for the proportion - calcuated directly using `qnorm()` (see page 18 in Gelman and Hill)
```{r}
d3 <- d0 %>% 
 filter(edattain != 0,
         edattain != 9) %>%
  mutate(edattain = as.factor(edattain)) %>%
  group_by(edattain) %>%
  summarise(male = sum(sex == 1),
            female = sum(sex == 2),
            n = male + female,
            p = female/n ,
            se = sqrt(p* (1-p) / n),  ###v
            lwr = p + qnorm(p = 0.025) * se,
            upp = p + qnorm(p = 0.975) * se)
```

18. Plot the confidence intervals for each proportion in `d3`. Include a horizontal line at 0.5.
```{r}
ggplot(data = d3, 
       mapping = aes(x = edattain, y = p)) +
  geom_point() +
  geom_pointrange(mapping = aes(ymin = lwr, ymax = upp)) +
  geom_line(yintercept = 0.5)
```

19. Q: What can we say about the confidence intervals?


A: ##### 