---
title: "Week 3 Exercise 1: IPUMS Simple Linear Regression"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### A. Set Up

1. Load the `haven`, `tidyverse` and `ipumsr` packages

```{r, message=FALSE}
library(haven)
library(tidyverse)
library(ipumsr)
```

2. Read in the `ipumsi_pan1960.dta` file, filter to those in Panama City (geo code for admin district 1 of 591008), save as object `d0`

```{r}
d0 <- read_dta("C:/Users/18800/Desktop/stat model/data/ipumsi_pan1960.dta") %>%
  filter(geolev1==591008)

d0
```

#### Obtaining Marital Age Difference

3. Create object `d1` that 

    a. contains only those with a spouse in the household
    b. `age` is numeric
    c. contains a new variable that uniquely identifies a married couple in a household
    d. arranged by household number, couple id and gender
    e. for each couple in each household summarise 
        i) the husband excess years, 
        ii) if the house is in an urban area (1 if yes, 0 if not)
        iii) husbands ages

```{r}
d1 <- d0 %>%
  filter(sploc !=0) %>%
 mutate(age = as.numeric(age), 
         couple_id = paste(pmin(pernum, sploc), pmax(pernum, sploc), sep = "-")) %>%
  arrange(serial, couple_id, sex) %>%
  group_by(serial, couple_id) %>%
  summarise(husband_older_by = age[1] - age[2],
            urban_house = as.numeric(urban[1]) - 1,
            husband_age = age[1])
d1
```


#### Averages with regression models

4. Find the average age that the husband is older than the wife using
    
    a. the mean function and data `d1`
    b. fitting a regession model with only a constant term

```{r}
mean(d1$husband_older_by)
lm(formula = husband_older_by ~ 1, data = d1)
```

5. Find the differences in the average excess of husbands ages in urban and rural areas using a regression model

```{r}
m0 <- lm(formula = husband_older_by ~ urban_house, data = d1)
m0
```
Q: What does the intercept term represent. 
A: in rural area, the men is usually 5.5142 years older than women 

Q: What does the parameter estimate for the urban represent
A: in urban area,the marriage age difference is 0.7511 smaller than that in urban area

6. Plot the scatter plot from `d1` that shows the 
  
    a. observations of urban status against the husband age excess with points jittered with width 0.05 so that overlapping obeservations can be viewed
    b. the fit of a linear model between the two variables
    

```{r}
ggplot(data = d1,
       mapping = aes(x = urban_house, y = husband_older_by)) +
  geom_point(position = position_jitter(width = 0.05)) +
  geom_smooth(method='lm') 
```

#### Modelling Martial Age Differences

7. Use LaTeX to write an equation of the regression model for the outcome variable $y_i$ regressed on a constant term $\beta_0$ and predictor variable $X$_i 

$$
y_i = \beta_0 + x_i + \epsilon_i
$$
    

8. Run a regression model that 
    - predicts the age difference based on the male age, 
    - saves the model object as `m1`

```{r}
m1 <- lm(formula = husband_older_by ~ age[1], data = d1)
m1
```

9. Using the results from `m1`, write an equation based on the fitted values for $\beta_0$ and $beta_1$.
$$
\texttt{age_difference} =  \beta_0  + \beta_1 \times {husband\_age} + \texttt{error}
$$
Q: Interpret the estimates of $\beta_0$ and $beta_1$.
A: #####

10. Plot the scatter plot from `d1` that shows the 
    
    a. observations of husband age against the husband age excess
    b. the fit of a linear model between the two variables

```{r}
ggplot(data = d1,
       mapping = aes(x = husband_age, y = husband_older_by)) +
  geom_point() +
  geom_smooth(method='lm')
  
```


#### Multiple Linear Regression 


11. Run a regression model that 
    - predicts the age difference based on the urban status and husband age, 
    - saves the model object as `m2` 
```{r}
m2 <- lm(formula = husband_older_by ~ urban_house + husband_age, data = d1)
m2
```

Q: Write an interpretation each of the coefficients
A: #####


12. Use the `ggpredict()` function to predict the values based on model `m2` where both terms are in the model. Use the default plot to show the marginal effects
```{r}
# install.packages("ggeffects")
library(ggeffects)
e2 <- ggpredict(model = m2, terms = c("husband_age", "urban_house"))
plot(e2)
```

13. Add the raw data to the plot using
    a) default ggpredict method
    b) your own ggplot code with a reduced alpha
    
```{r, eval=FALSE}
# gives an error... need to make urban_hose a factor
plot(e2, rawdata = TRUE)
```

```{r}
lm(formula = husband_older_by ~ factor(urban_house) + husband_age, data = d1) %>%
  ggpredict(terms = c("husband_age", "urban_house")) %>%
  plot(rawdata = TRUE)


# more flexible ggplot method
ggplot(data = e2, mapping = aes(x = x, y = predicted, colour = group)) + 
  geom_line() +
  geom_ribbon(mapping = aes(ymin = conf.low, ymax = conf.high, fill = group), 
              alpha = 0.25, linetype = 1) +
  geom_point(data = d1, 
             mapping = aes(x = husband_age, y = husband_older_by, colour = factor(urban_house)),
             alpha = 0.1)
```    

